---
title: "Using the lasso estimator with regmodelsuite"
output: 
  rmarkdown::html_vignette:
    toc: true          
    toc_depth: 2       
vignette: >
  %\VignetteIndexEntry{Using the lasso estimator with regmodelsuite}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Generate Test Data

This vignette demonstrates how to estimate a LASSO regression model using the `regmodelsuite` package. LASSO is a type of linear regression that includes a penalty to shrink some coefficients to zero, effectively performing variable selection.

First we are going to create some example Data where the number of predictors
is greater than the number of observations p > n

```{r}
library(regmodelsuite)

set.seed(1443)  # For reproducibility

# Parameters
n <- 50  # Number of observations
p <- 100  # Number of predictors

# Generate predictor matrix (X) with normal distributed values
X <- matrix(rnorm(n * p), nrow = n, ncol = p)

# Create a true coefficient vector
beta_true <- rnorm(p)

# zero out some coefficients
beta_true[abs(beta_true) < 0.5] <- 0

# Generate response variable (y) as a linear combination of predictors + noise
y <- X %*% beta_true + rnorm(n)

# Convert to data frame
full_data <- data.frame(X)
full_data$y <- y


# Split the dataset into a training and test set (50/50)

training_indices <- sample(1:nrow(full_data), size = 0.5 * floor(nrow(full_data)))

training_set <- full_data[training_indices , ]
test_set <- full_data[-training_indices , ]

cat("Full data dimensions: ", dim(full_data), "\n")
cat("Training set dimensions: ", dim(training_set), "\n")
cat("Test set dimensions: ", dim(test_set), "\n")

```
## Estimate a lasso model with cross validation

To estimate a lasso model with cross validation the user has to supply a formula 
into the regmodel function. We are going to add all p predictors.

Here we let the \code{regmodel} function create a lambda grid with 10 lambdas
with 5 folds for cross validation.
```{r, results = "hide", message = FALSE, warning = FALSE}
set.seed(2302)

model_formula <- y ~ .

lassoCV <- regmodel(model_formula, data = training_set, model = "lasso", cv = T,
                    nlambda = 10, m = 5)


```
\ 
```{r}
print(lassoCV)
```



## Specify a user defined lambda grid

It is also possible to specify a user defined lambda grid.
```{r}
lambda_grid <- runif(10 , min = 0,  max = lassoCV$min_lambda + 2)
print(lambda_grid)
```
\ 
```{r, results = "hide", message = FALSE, warning = FALSE}

lassoCVmanual <- regmodel(model_formula , data = training_set, model = "lasso",
                          cv = T, lambda = lambda_grid, m = 5)
```
\ 
```{r}
print(lassoCVmanual)
```

## Estimating the final model

The \texttt{lassoCVmanual} object contains information about the cross validation 
process. To use the estimated lambda for the final model we can use
\code{regmodel} again.It is also possible to extract scaled and unscaled 
coefficients from the model by using \code{coef}.

```{r}
lasso_fit <- regmodel(model_formula,
                      data = training_set,
                      model = "lasso",
                      lambda = lassoCVmanual$min_lambda)

# scaled coefficients
coef(lasso_fit)
```
\ 
```{r}
# unscaled coefficients
coef(lasso_fit, unscale = TRUE)
```


## Predictions 

Predictions can be made by using the predict function.
Just calling predict on a lasso object will use the training data set for 
predictions. By specifying newdata, the estimated model will be applied to 
the new data.

```{r, fig.width=6, fig.height=4}
pred_insample <- predict(lasso_fit)
print(pred_insample)

# Now use the test set
pred_outofsample <- predict(lasso_fit, newdata = test_set)
print(pred_outofsample)

# Calculate the Mean squared prediction error
mean((pred_outofsample - test_set$y)^2)

# Plot y and predicted y
{
 plot(test_set$y);
 points(pred_outofsample, col = "red" , pch = 2)
}
```

We can visualize the Mean squared prediction error
```{r, fig.width=6, fig.height=4}
{
  plot(log(lassoCV$lambda_grid), lassoCV$MSPE,
     xlab = "Log(Lambda)", ylab = "Mean Squared Prediction Error",
     main = "MSPE vs. Log(Lambda)")
}
```

