---
title: "Using the lasso estimator with regmodelsuite"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lasso}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This vignette demonstrates how to estimate a LASSO regression model using the `regmodelsuite` package. LASSO is a type of linear regression that includes a penalty to shrink some coefficients to zero, effectively performing variable selection.



```{r}
library(Ecdat)
library(dplyr)
library(regmodelsuite)

df <- Caschool %>%
  select(-district)
head(df)

# Split the data into a Training Set and Test Set (60/40)
indexes <- sample(1:(nrow(df) * 0.5))
training_set <- df[indexes , ]
test_set <- df[-indexes, ] 

dim(df)
dim(training_set)
dim(test_set)

```
In order to estimate a lasso model the user can supply a formula into the regmodel function. We are going to use the amount of cyclinders, horsepower, weight and transmission type for predicting miles per gallon. The variables "cyl" and "am" will be included as dummy variables.
First we are going to calculate value for lambda by using cross-validation

```{r}
# seed for reproducebillity
set.seed(2302)
model_formula <- mathscr ~ .*.

ridgeCV <- regmodel(model_formula, data = training_set, model = "ridge", cv = T)

min_lambda <- ridgeCV$min_lambda

print(ridgeCV)


```
The "lasso_fit" object contains information about the cross validation process. To use the estimated lambda for the final model we can use \code{regmodel} again

```{r}
lasso_fit <- regmodel(model_formula,
                      data = training_set,
                      model = "lasso",
                      lambda = min_lambda)

print(lasso_fit)
```
In this case three variables were selected by the lasso model

Predictions can be done by using the \code{predict()} function.
Just calling predict on a lasso object will use the training data set for 
predictions. By specifying \code{newdata}, the estimated model will be applied to 
the new data.

```{r}
pred_insample <- predict(lasso_fit)
print(pred_insample)

# Now use the test set
pred_outofsample <- predict(lasso_fit, newdata = test_set)
```

We can visualize the predictions using ggplot2
```{r}

```

