---
title: "Stepwise Selection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Stepwise Selection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(regmodelsuite)
```

## Data
For the stepwise selection examples, we will use the built-in `mtcars` dataset:
```{r}
head(mtcars)
```

## Running Stepwise Selection
To run a stepwise selection, you need to call the wrapper function 
`regmodel()` with the `model` parameter set to `"forward"` for forward
selection or `"backward"` for backward selection. You can then set the
`formula` parameter to a formula object and the `data` parameter to a
data frame. To specify the number of predictors to select, you need to set
the `n_predictors` parameter. Here, we select 2 predictors using forward
selection for a model predicting `mpg` from all the other columns:
```{r}
forward <- regmodel(formula = mpg ~ ., data = mtcars, model = "forward",
                    n_predictors = 2)
```
To understand how the predictors were selected, you can check what the function
printed. It first tried each predictor on its own, then selected `wt`
because it had the lowest average error. It then tried each predictor again,
combined with `wt`, and selected `cyl` because it had the lowest
average error.  

If you do not want the function to print this, you can set the `verbose`
parameter to `FALSE`:
```{r}
forward <- regmodel(formula = mpg ~ ., data = mtcars, model = "forward",
                    n_predictors = 5, verbose = FALSE)
```

You can run backward selection in the same way.
Here, we select 8 predictors using backward selection for the same model:
```{r}
backward <- regmodel(formula = mpg ~ ., data = mtcars, model = "backward",
                     n_predictors = 8)
```
Again, you can see how the predictors were selected by checking what the
function printed. It first tried all predictors and then removed `cyl`
because it had the lowest contribution to the model, measured by the Z-score.
It then tried all predictors except `cyl` and removed `vs` because it had
the lowest Z-score. All the other predictors were therefore selected.

## Using the stepwise selection object
After running a stepwise selection, you get an S3 object of the type
`stepwise_selection`. To see the result of the selection, you can simply print
the object:
```{r}
forward
```
You can see which predictors were selected, what the error of the final model
has, and in which direction the selection was made.  
To get the selected parameters as a character vector, you can use the
`predictors` attribute:
```{r}
forward$predictors
```

If you want to predict the data with the model, you can either extract the model
using the `model` attribute or use the `predict` function directly on the
stepwise selection object:
```{r}
predict(forward)
```

If you want to predict new data, you can supply data of the same format using
the `new_data` parameter. Here, we use an edited version of `mtcars`:
```{r}
new_data <- mtcars * 0.5
predict(forward, new_data[forward$predictors])
```

## Using Different Models
By default, stepwise selection uses the built-in `lm` regression model, but
you can also use different models by supplying a function with the `model_fct`
parameter. This function must have a `data` parameter that accepts a
data frame and a `formula` parameter that accepts a formula object. The
function must also return an object with an implemented `predict` function.
For example, you can use the built-in `glm` function:
```{r}
regmodel(formula = mpg ~ ., data = mtcars, model = "forward",
         n_predictors = 2, model_fct = glm, verbose = FALSE)
```

You can also use the models implemented in this package, but you need to put
the function call inside another function. For example, this is how you can use
ridge regression:
```{r}
regmodel(formula = mpg ~ ., data = mtcars, model = "forward",
         n_predictors = 2, verbose = FALSE,
         model_fct=\(formula, data) regmodel(formula, data, "ridge", lambda=0))
```

## Visualizing the Selection
To visualize the selection process of a stepwise selection object as a graph,
you can use the `plot` function. It shows the error at each step,
with every number of predictors until the specified amount is reached:
```{r}
plot(forward)
```

You can also get the values of the errors as a vector if you want to visualize
them yourself:
```{r}
forward$errors
```
